---
title: "IST687 Project Master Code"
author: "Megha Banerjee, Kevin Harmer"
date: "12/9/2021"
output:
  pdf_document: default
  html_document: default
---


1. Collective Libraries Used for Analysis
```{r}
library(tidyverse)
library(ggplot2)
library(stats)
library(car)
library(caret)
library(kernlab)
library(e1071)
library(rworldmap)
library(arules)
library(arulesViz)
library(performance)
library(DHARMa)
library(stats)
library(dplyr)
library(purrr)
library(tidyr)
```

2. Importing Data
```{r}
hotels <- read_csv('https://intro-datascience.s3.us-east-2.amazonaws.com/Resort01.csv') #importing the data
```

3. Checking for Any NA values and Cleaning
```{r}
#View(hotels) #viewing data set in other window
str(hotels) #view data set structure
head(hotels) #view the first 5 inputs
for (i in c(1:20)) {
  print(sum(hotels[,i] == 'NULL')) #checking for NULL inputs; also checked for NA values but there were not any
} #amount of null values in data set
hotels <- hotels[!hotels$Country == 'NULL',] #removing inputs without a country input
```

```{r}
cancel <- hotels[hotels$IsCanceled==1,] #creating a new data set with only canceled observations
nrow(cancel) #looking at amount of data observations

stay <- hotels[hotels$IsCanceled==0,] #creating a new data set with only non-canceled observations
nrow(stay) #looking at amount of dat observations
```

4. Looking at Some Bar Plots
```{r}
# COLUMN NOT REQUIRED for analysis of cancellation, no distinct measure shows up 
plotyplot <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$StaysInWeekendNights)
plotyplot
# COLUMN NOT REQUIRED for analysis of cancellation, no distinct measure shows up 
plotyplot_2 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$LeadTime)
plotyplot_2
# COLUMN NOT REQUIRED for analysis of cancellation, no distinct measure shows up 
plotyplot_3 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$StaysInWeekNights)
plotyplot_3
# bookings with 2 adults often tend to get cancelled more frequently , around 75 percent of the times in respect to bookings made where there were 1 or 3 adults
plotyplot_4 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$Adults)
plotyplot_4
# Guests who have children hardly cancel but there cancel rate is higher than with babies
plotyplot_5 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$Children)
plotyplot_5
# Guests who have babies hardly cancel.
plotyplot_6 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$Babies)
plotyplot_6
# only 1 percent of the repeated guests cancel on the hotels. 
plotyplot_7 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$IsRepeatedGuest)
plotyplot_7
# People with Meal SC and FB hardly cancel whereas people taking BB meals cancel 80 percent of the times
plotyplot_8 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$Meal)
plotyplot_8
# COLUMN NOT REQUIRED for analysis

# there is only one country from where bookings are cancelled the most. While we can also infer that most of the bookings are made from this country. Thus this is not an effective measure to find out the probability of cancellation. 
plotyplot_9 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x=cancel$Country)
plotyplot_9
# the market segment which comprises of online TA is responsible for most of the cancellations. Bookings made through groups are second is this respect
plotyplot_10 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$MarketSegment)
plotyplot_10
# COLUMN NOT REQUIRED for analysis

# We cannot use this as a measure to identify potential risk of someone cancelling their booking. Since 99 percent of people had never cancelled their bookings in the hotels before
plotyplot_11 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$PreviousCancellations)
plotyplot_11
# COLUMN NOT REQUIRED for analysis

# We cannot use this as a measure to identify potential risk of someone cancelling their booking. Since 99 percent of people had never cancelled their bookings in the hotels before
plotyplot_12 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$PreviousBookingsNotCanceled)
plotyplot_12
# people assigned with room type A has seen most of the cancellations. To avoid cancellations , people who are identified as high risk cancel probability should be assigned room type B if possible and not A
plotyplot_13 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$ReservedRoomType)
plotyplot_13
# people assigned with room type A has seen most of the cancellations. To avoid cancellations , people who are identified as high risk cancel probability should be assigned room type B if possible and not A
plotyplot_14 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$AssignedRoomType)
plotyplot_14
# most people directly cancel there plans. But some have to change their bookings to accomodate their plan to stay in the hotels. If that is not possible then they cancel their bookings
plotyplot_15 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$BookingChanges)
plotyplot_15
# when booking made with no deposit
plotyplot_16 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$DepositType)
plotyplot_16
# transient customers are the types who are cancelling the booking the most
plotyplot_17 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$CustomerType)
plotyplot_17
# people who cancelled their bookings do not ask for a parking space in advance.
plotyplot_18 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$RequiredCarParkingSpaces)
plotyplot_18
# people with 0 special requests tend to cancel their bookings. We should be sure the booking won't be cancelled if the number of special requests is more than 4
plotyplot_19 <-  ggplot(cancel) + geom_bar(data=cancel)+ aes(x= cancel$TotalOfSpecialRequests)
plotyplot_19
```

Actionable Insights:
1) While selecting Room type either reserved or assigned room type , avoid giving room type A to people show  patterns of high cancellations risk.
2) Try incorporating a structure where deposit amount is compulsory because most of the people cancelled had no deposit
3) When people make an effort to change their bookings greet them with more importance because it shows inclination towards coming to stay in the hotels. Which means there is less chance of cancellation
4) Make an effort to get special requests from the customers, to be very safe , at least 4 or more. Customers making special requests hardly cancel their bookings
5) Always ask for parking space from the customer prior to the booking. Reserving parking space for the customers will lead to reduction in cancellation.
6) The above 5 points should be focused especially more on the market segment "Online TA" which is responsible for most of the cancellations
7) Special care of the first 5 points should be given to transient customers as they cancel more than other type of customers
8) Customers who are families especially who have children and babies would hardly cancel bookings in your hotels, focus on customers in groups of 2.


5. Geogrphic Analysis
```{r}
x<- hotels %>% #filtering hotel data set
  group_by(Country) %>% #by countries
  summarise(Frequency = sum(as.numeric(IsCanceled))) #and summing 

```

```{r}
joinData <- joinCountryData2Map( x, #joining our country subset
                                 joinCode = "ISO3", #with code
                                 nameJoinColumn = "Country") #to countries
```

```{r}
#WorldMap
theMap <- mapCountryData( joinData, nameColumnToPlot="Frequency", mapTitle= 'World Map Cancellations Frequency', addLegend=TRUE,catMethod = "Pretty", colourPalette = "Terrain", oceanCol = 'light blue',missingCountryCol = 'transparent')
#do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))
```
```{r}
#EUROPE
theMap <- mapCountryData( joinData, nameColumnToPlot="Frequency", mapTitle= 'Europe Cancellations Frequency',addLegend=TRUE,catMethod = "Pretty", colourPalette = "Terrain", oceanCol = 'light blue',missingCountryCol = 'transparent',mapRegion = "Europe" )
#do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))

#country_coord<-data.frame(coordinates(joinData),stringsAsFactors=F)
# label the countries
#text(x=country_coord$X1,y=country_coord$X2,labels=row.names(country_coord))
```


```{r}
#Africa
theMap <- mapCountryData( joinData, nameColumnToPlot="Frequency", mapTitle= 'Africa Cancellations Frequency',addLegend=TRUE,catMethod = "Pretty", colourPalette = "Terrain", oceanCol = 'light blue',missingCountryCol = 'transparent',mapRegion = "Africa" )
#do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))

#country_coord<-data.frame(coordinates(joinData),stringsAsFactors=F)
# label the countries
#text(x=country_coord$X1,y=country_coord$X2,labels=row.names(country_coord))
```


```{r}
#ASIA

theMap <- mapCountryData( joinData, nameColumnToPlot="Frequency", mapTitle= 'Asia Cancellations Frequency',addLegend=TRUE,catMethod = "Pretty", colourPalette = "Terrain", oceanCol = 'light blue',missingCountryCol = 'transparent',mapRegion = "Asia" )
#do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))

#country_coord<-data.frame(coordinates(joinData),stringsAsFactors=F)
# label the countries
#text(x=country_coord$X1,y=country_coord$X2,labels=row.names(country_coord))
```


```{r}
#Europe-Asia

theMap <- mapCountryData( joinData, nameColumnToPlot="Frequency", mapTitle= 'Eurasia Cancellations Frequency',addLegend=TRUE,catMethod = "Pretty", colourPalette = "Terrain", oceanCol = 'light blue',missingCountryCol = 'transparent',mapRegion = "eurasia" )
#do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))

#country_coord<-data.frame(coordinates(joinData),stringsAsFactors=F)
# label the countries
#text(x=country_coord$X1,y=country_coord$X2,labels=row.names(country_coord))
```


```{r}
#Latin america

theMap <- mapCountryData( joinData, nameColumnToPlot="Frequency", mapTitle= 'Latin America Cancellations Frequency',addLegend=TRUE,catMethod = "Pretty", colourPalette = "Terrain", oceanCol = 'light blue',missingCountryCol = 'transparent',mapRegion = "latin america" )
#do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))

#country_coord<-data.frame(coordinates(joinData),stringsAsFactors=F)
# label the countries
#text(x=country_coord$X1,y=country_coord$X2,labels=row.names(country_coord))
```


```{r}
#North america

theMap <- mapCountryData( joinData, nameColumnToPlot="Frequency", mapTitle= 'North America Cancellations Frequency',addLegend=TRUE,catMethod = "Pretty", colourPalette = "Terrain", oceanCol = 'light blue',missingCountryCol = 'transparent',mapRegion = "north america" )
#do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))

#country_coord<-data.frame(coordinates(joinData),stringsAsFactors=F)
# label the countries
#text(x=country_coord$X1,y=country_coord$X2,labels=row.names(country_coord))
```



```{r}
#Oceania

theMap <- mapCountryData( joinData, nameColumnToPlot="Frequency", mapTitle= 'Oceania Cancellations Frequency',addLegend=TRUE,catMethod = "Pretty", colourPalette = "Terrain", oceanCol = 'light blue',missingCountryCol = 'transparent',mapRegion = "oceania" )
#do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))

#country_coord<-data.frame(coordinates(joinData),stringsAsFactors=F)
# label the countries
#text(x=country_coord$X1,y=country_coord$X2,labels=row.names(country_coord))
```



```{r}
#United Kingdom

theMap <- mapCountryData( joinData, nameColumnToPlot="Frequency", mapTitle= 'United Kingdom Cancellations Frequency',addLegend=TRUE,catMethod = "Pretty", colourPalette = "Terrain", oceanCol = 'light blue',missingCountryCol = 'transparent',mapRegion = "uk" )
#do.call( addMapLegend, c(theMap, legendWidth=1, legendMar = 2))

#country_coord<-data.frame(coordinates(joinData),stringsAsFactors=F)
# label the countries
#text(x=country_coord$X1,y=country_coord$X2,labels=row.names(country_coord))
```

6. Miscellaneous Graphing
```{r}
ggplot(hotels, aes(factor(IsCanceled), #bar plot of only the IsCanceled variable
        fill = factor(IsCanceled))) +geom_bar()+ xlab("0= Non-Cancelled booking, 1= Cancelled booking")+ ylab("Count")

ggplot(hotels, aes(factor(IsRepeatedGuest), #bar plot describing the IsRepeatedGuest variable with IsCanceled fill
        fill = factor(IsCanceled))) +geom_bar() + xlab("Repeated Guest(1) or not (0) \n 0= Non-Cancelled booking, 1= Cancelled booking")+ ylab("Count")

ggplot(hotels, aes(factor(MarketSegment), #bar plot describing the MarketSegment variable with IsCanceled fill
        fill = factor(IsCanceled))) +geom_bar()+ xlab("Market Segment\n 0= Not-Cancelled booking, 1= Cancelled booking")+ ylab("Count")

ggplot(hotels, aes(factor(DepositType), #bar plot describing the DepositType variable with IsCanceled fill
        fill = factor(IsCanceled))) +geom_bar()+ xlab("Deposit Type\n 0= Non-Cancelled booking, 1= Cancelled booking")+ ylab("Count")

ggplot(hotels, aes(factor(CustomerType), #bar plot describing the CustomerType variable with IsCanceled fill
        fill = factor(IsCanceled))) +geom_bar()+ xlab("Customer Type\n 0= Non-Cancelled booking, 1= Cancelled booking")+ ylab("Count")
```


```{r}
hotels %>% #generalized bar plots with each of the variables
  keep(is.numeric) %>% #changing factors to numeric where possible
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_bar()
```

7. Descriptive Statistics and Associated Plots
```{r}
numdata <- hotels[,c('IsCanceled','LeadTime','StaysInWeekendNights','StaysInWeekNights',
                     'PreviousCancellations','PreviousBookingsNotCanceled','BookingChanges',
                     'RequiredCarParkingSpaces')] #subsetting only the numerical columns
numdata$IsCanceled <- factor(numdata$IsCanceled, levels = c(0,1), labels = c('Not Canceled', 'Canceled')) #changing IsCanceled to Factored Data
newstay <- numdata[numdata$IsCanceled == 'Not Canceled',][,-1] #data set with all of the non-cancellations
newcancel <- numdata[numdata$IsCanceled == 'Canceled',][,-1] #data set with all of the cancellations
summary(newstay) #summary statistics of stay observations
summary(newcancel) #summary statistics of cancel observations

numdata$IsCanceled <- as.numeric(numdata$IsCanceled)-1 #bringing back to numeric only for the next function
cor(numdata) #correlation matrix between all of the numeric variables and IsCanceled as numeric
numdata$IsCanceled <- factor(numdata$IsCanceled, levels = c(0,1), labels = c('Not Canceled', 'Canceled')) #changing IsCanceled back to Factored Data
```

Descriptive Statistics to note:
- Lead time appears much larger in cancellation data set

- IsRepeatedGuest has a significantly higher mean in stays

- PreviousCancellations has a significantly higher mean in canceled

- PreviousBookingsNotCanceled has a significantly higher mean in stays

- BookingChanges has a higher mean in stays

```{r}
boxplot(LeadTime ~ IsCanceled, data = numdata) #LeadTime
boxplot(StaysInWeekendNights ~ IsCanceled, data = numdata) #StaysInWeekendNights
boxplot(StaysInWeekNights ~ IsCanceled, data = numdata) #StaysInWeekNights
boxplot(PreviousCancellations ~ IsCanceled, data = numdata) #PreviousCancellations
boxplot(PreviousBookingsNotCanceled ~ IsCanceled, data = numdata) #PreviousBookingsNotCanceled
boxplot(BookingChanges ~ IsCanceled, data = numdata) #BookingChanges
boxplot(RequiredCarParkingSpaces ~ IsCanceled, data = numdata) #RequiredParkingSpaces
```

Modified box plots to Get a Better Idea of Variable Distribution (temporily remove outliers 3 standard deviations away from mean)
```{r}
boxplot(LeadTime ~ IsCanceled, data = numdata[numdata$LeadTime<mean(numdata$LeadTime)+3*sd(numdata$LeadTime),]) #LeadTime
boxplot(StaysInWeekendNights ~ IsCanceled, data = numdata[numdata$StaysInWeekendNights < mean(numdata$StaysInWeekendNights)+
                                                            3*sd(numdata$StaysInWeekendNights),]) #StaysInWeekendNights
boxplot(StaysInWeekNights ~ IsCanceled, data = numdata[numdata$StaysInWeekNights < mean(numdata$StaysInWeekNights)+
                                                         3*sd(numdata$StaysInWeekNights),]) #StaysInWeekNights
boxplot(PreviousCancellations ~ IsCanceled, data = numdata[numdata$PreviousCancellations < mean(numdata$PreviousCancellations)+
                                                             3*sd(numdata$PreviousCancellations),]) #PreviousCancellations
boxplot(PreviousBookingsNotCanceled ~ IsCanceled, data = numdata[numdata$PreviousBookingsNotCanceled < mean(numdata$PreviousBookingsNotCanceled)+
                                                                   3*sd(numdata$PreviousBookingsNotCanceled),]) #PreviousBookingsNotCanceled
boxplot(BookingChanges ~ IsCanceled, data = numdata[numdata$BookingChanges < mean(numdata$BookingChanges)+
                                                      3*sd(numdata$BookingChanges),]) #BookingChanges
boxplot(RequiredCarParkingSpaces ~ IsCanceled, data = numdata[numdata$RequiredCarParkingSpaces <
                                                                mean(numdata$RequiredCarParkingSpaces)+
                                                                3*sd(numdata$RequiredCarParkingSpaces),]) #RequiredParkingSpaces
```

Looking at Boxplots of LeadTime in Correspondance with other factors

```{r}
stayfull <- hotels[hotels$IsCanceled == 0,][,-20][,-1] #data set with all of the non-cancellations
cancelfull <- hotels[hotels$IsCanceled == 1,][,-20][,-1] #data set with all of the cancellations
```

```{r Adult_Boxplots}
boxplot(LeadTime ~ Adults, data = stayfull) #stays
boxplot(LeadTime ~ Adults, data = cancelfull) #cancels
```

8. Running some models
```{r}
linmod <- lm(as.numeric(IsCanceled) ~ ., data = numdata) #running a linear model. Won't be super accurate because IsCanceled is dichotomous; need log
summary(linmod)

logmod <- glm(IsCanceled ~ ., family = binomial(link="logit"), data = numdata[-13920,]) #major outlier at 13920; 27 previous bookings attended
summary(logmod)

logmod2 <- glm(IsCanceled ~ LeadTime, family = binomial(link="logit"), data = numdata) #just checking out Lead Time
summary(logmod2)

model_performance(logmod) #used to find psuedo r-squared of logistics
anova(logmod, test = "Chisq") #analyzing chi-squared results

#vif(linmod) #no colinarity!
#vif(logmod) #none here either
```

Predictions:
```{r}
linout <- predict(linmod, numdata[,-1]) #prediction data with linear model
linout <- data.frame(linout) #merging it into data frame
linout$data <- 0 #creating new column of 0's
#linout$data[linout$linout > as.numeric(median(linout$linout))] <- 1
linout$data[linout$linout > 1.5] <- 1 #1.5 is the halfway point in the factored IsCanceled data; make the top half 1
linout$data <- factor(linout$data, levels = c(0,1), labels = c('Not Canceled', 'Canceled')) #changing new column to Factored Data

confusionMatrix(as.factor(linout$data),numdata$IsCanceled) #creating a confusion matrix to analyze results

logout <- predict(logmod, numdata[,-1]) #prediction data with logistic model
logout <- data.frame(logout) #merging it into data frame
logout$data <- 0 #creating new column of 0's
#logout$data[logout$logout > as.numeric(median(logout$logout))] <- 1
logout$data[logout$logout > 1.5] <- 1 #1.5 is halfway point in the factored IsCanceled data; make the top half 1
logout$data <- factor(logout$data, levels = c(0,1), labels = c('Not Canceled', 'Canceled')) #changing new column to Factored Data

confusionMatrix(as.factor(logout$data),numdata$IsCanceled) #creating confusion matrix to analyze results
```

Analytical Plots
```{r}
plot(x = numdata$LeadTime, y = numdata$IsCanceled) #ploting canceled vs leadtime
plot(linmod) #plotting linear model residuals
```

```{r}
simulationOutput <- simulateResiduals(fittedModel = logmod, n = 250) #simulaiing the logarithmic residuals
plot(simulationOutput) #ploting the simulation

plot(logmod) #outlier at 13920; removed to better view graphic
```

Looking at Variable Interactions
```{r}
linintmod <- lm(as.numeric(IsCanceled) ~ .+StaysInWeekendNights:StaysInWeekNights+PreviousCancellations:PreviousBookingsNotCanceled, data = numdata) #running a linear model. Won't be super accurate because IsCanceled is dichotomous; need log
summary(linintmod)

logintmod <- glm(IsCanceled ~  .+StaysInWeekendNights:StaysInWeekNights, family = binomial(link="logit"), data = numdata[-13920,]) #major outlier at 13920; 27 previous bookings attended
summary(logintmod)

linintout <- predict(linintmod, numdata[,-1]) #interaction predictions; same method as before
linintout <- data.frame(linintout)
linintout$data <- 0
#linout$data[linout$linout > as.numeric(median(linout$linout))] <- 1
linintout$data[linintout$linintout > 1.5] <- 1 #1.5 is the halfway point in the factored IsCanceled data
linintout$data <- factor(linintout$data, levels = c(0,1), labels = c('Not Canceled', 'Canceled')) #changing new column to Factored Data

confusionMatrix(as.factor(linintout$data),numdata$IsCanceled)

logintout <- predict(logintmod, numdata[,-1]) #interaction predictions; same method as before
logintout <- data.frame(logintout)
logintout$data <- 0
#logout$data[logout$logout > as.numeric(median(logout$logout))] <- 1
logintout$data[logintout$logintout > 1.5] <- 1
logintout$data <- factor(logintout$data, levels = c(0,1), labels = c('Not Canceled', 'Canceled')) #changing new column to Factored Data

confusionMatrix(as.factor(logintout$data),numdata$IsCanceled)
```

9. Moving on to Factors
```{r}
#changing hotel categorical variables into factored variables
hotels$IsCanceled <- as.factor(hotels$IsCanceled)
hotels$Adults <- as.factor(hotels$Adults)
hotels$Children <- as.factor(hotels$Children)
hotels$Babies <- as.factor(hotels$Babies)
hotels$Meal <- as.factor(hotels$Meal)
hotels$Country <- as.factor(hotels$Country)
hotels$MarketSegment <- as.factor(hotels$MarketSegment)
hotels$IsRepeatedGuest <- as.factor(hotels$IsRepeatedGuest)
hotels$ReservedRoomType <- as.factor(hotels$ReservedRoomType)
hotels$AssignedRoomType <- as.factor(hotels$AssignedRoomType)
hotels$DepositType <- as.factor(hotels$DepositType)
hotels$CustomerType <- as.factor(hotels$CustomerType)

facdata <- data.frame(hotels[,-20][,-19][,-16][,-13][,-12][,-4][,-3][,-2], #subsetting the numerics and the columns changing into factors
                         PreviousCancellations = factor(hotels$PreviousCancellations > 0, levels = c(FALSE, TRUE), labels = c('None','Some')), #factoring PreviousCancellations
                         PreviousBookingsNotCanceled = factor(hotels$PreviousBookingsNotCanceled > 0, levels = c(FALSE, TRUE), labels = c('No Attended Resrvation','Have Followed Booking')), #factoring PreviousBookingsNotCanceled
                         BookingChanges = factor(hotels$BookingChanges > 0, levels = c(FALSE, TRUE), labels = c('No Booking Changes', 'Some Booking Changes')), #factoring BookingChanges
                         RequiredParkingSpaces = factor(hotels$RequiredCarParkingSpaces > 0, levels = c(FALSE, TRUE), labels = c('No Required Parking Spaces', 'Required Parking Space')), #factoring RequiredCarParkingSpaces
                         TotalOfSpecialRequests = factor(hotels$TotalOfSpecialRequests > 0, levels = c(FALSE, TRUE), labels = c('No Special Requests', 'Made Special Requests'))) #factoring number of special requests
```

10. SVM
```{r}
#svm part 1
set.seed(100) #setting random number generator
trainList <- createDataPartition(y=facdata$IsCanceled,p=.60,list=FALSE) #partitioning the data
summary(trainList) 
str(trainList)
#trainList
trainSet<- facdata[trainList, ] #making training data set with partitioned indices
testSet<- facdata[-trainList, ] #making testing data set with remaining observations
testSet<- data.frame(testSet) #converting test set into data frame
#str(trainSet)
#str(testSet)

#svm part2
svmModel<- ksvm(IsCanceled~., data=trainSet, C = 5, cross=3, prob.model=TRUE) #running svm model
svmModel

predOut<- predict(svmModel, testSet) #getting predictions from svm model using test data set

confusionMatrix(predOut,testSet$IsCanceled) #analyzing results in a confusion matrix
```

11. Association Rule Mining
```{r}
hotels$IsCanceled <- as.factor(hotels$IsCanceled) #repositioning IsCanceled to remain consistent with future code
hotelsX <- as(hotels, Class = 'transactions') #making a hotels data set with transactions
itemFrequencyPlot(hotelsX, topN = 10) #making an item frequency plot; not necessarily needed
```


```{r}
freq <- itemFrequency(hotelsX) #creating item frequency
str(freq)
items <- sort(freq) #sorting data in ascending :sort(-item)#Descending

head(items) # least bought items
tail(items)
```

```{r}
hotel_rule <- apriori(hotelsX, #apriori transactions
parameter=list(supp=0.008, conf=0.9), #0.8 \% of instances with 90% confidence
control=list(verbose=F),
appearance=list(default="lhs",rhs=("IsCanceled=1"))) #looking for canceled reservations
```

```{r}
inspectDT(hotel_rule) #inspecting rule
```
